### 2025-09-23 paper/sec/introduction.tex
First pass over integrating use cases better into section 1,.

\figref{scigen-example-website}, generated from our implementation, illustrates these two features. The upper
section shows an excerpt from a research paper comparing text encoding techniques~\citep{zhang18}. When a
reader hovers over the phrase \textfrag{does not further improve}, the corresponding data entries are
highlighted in blue on the left. Other fragments (e.g. \textfrag{better than}, \textfrag{further
improvements}) that refer to the same data are also marked, allowing readers to explore supporting and
contrasting evidence. The lower section shows a counterfactual situation where we imagine the authors'
experiments had produced different results: here the phrase \textfrag{does not further improve} is replaced by
\textfrag{further improves}. Transparent documents thus turn static text into interactable, data-driven
content able to expose the evidential basis of scholarly claims.

@@

\begin{itemize}
\item a proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
associated data set into a transparent, data-driven counterpart
\item an empirical evaluation of how well state-of-the-art models are able to solve the associated
interpretation and code synthesis problems
\end{itemize}

@@

We envisage two possible use cases for transparent text:

\begin{enumerate}
\item \textbf{Authoring.} Someone authoring content for an online article, wants to create text
linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
the claims made in the text can be explored \emph{in situ}, by interacting with the text.

\item \textbf{Reading or reviewing.} Someone reading textual claims derived from open data (e.g. a
scientific paper or climate report), wants to retroactively link the text to queries over the available data
and gradually ``rationally reconstruct'' the relationship between the claims in the paper and the evidence
base. Perhaps just to aid their own comprehension, or to provide some kind of justified peer review.
\end{enumerate}

>>>>

\figref{scigen-example-website}, generated from our implementation, illustrates these two features. The upper
section shows an excerpt from \cite{zhang18}, a scholarly article comparing text encoding techniques. When a
reader hovers over the phrase \textfrag{does not further improve}, the relevant data are highlighted on the
left. Other fragments (e.g. \textfrag{better than}, \textfrag{further improvements}) that refer to the same
data are also marked, allowing the reader to explore supporting and contrasting evidence. The lower section
shows a counterfactual situation where the authors' experiments had produced different results: here the
phrase \textfrag{does not further improve} is replaced by \textfrag{further improves}.

Transparent documents thus turn static text into interactable, data-driven content able to expose the
evidential basis of scholarly claims, for which we envisage two use cases. First, when \textbf{authoring}
content for an online article, a journalist or scientific publisher may wish to link text to raw data so that
the evidence base for the textual claims can be explored directly from the article. Second, when
\textbf{reading} a document reporting on findings derived from open data (perhaps a scientific paper or
climate report), the reader may want to retroactively interpret parts of the text as queries over the
available data and gradually ``rationally reconstruct'' the relationship between claims in the paper and the
evidence base. This might be just to aid their own comprehension, or part of a formal peer review process.

@@

\begin{itemize}
\item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
associated data set into a transparent, data-driven counterpart;
\item An empirical evaluation of how well state-of-the-art models are able to solve the associated
interpretation and code synthesis problems.
\end{itemize}

@@
